Global Optimization of an Analog Method by Means of Genetic Algorithms

### 2016-09-02 Genetic algorithms - MWR/template.tex
GAs paper under correction.

In order to evaluate a combination of operators/options, 10 optimizations per configuration of GAs were processed. Such assessment is not possible on the whole archive length, and had to be performed on a reduced period. The performances were characterized by four indicators: (i) mean performance score: average of the final scores of the 10 optimizations, (ii) convergence: the number of optimizations that converged to a supposed global optimum, (iii) number of generations: characterization of the convergence speed, and (iv) number of evaluations of the objective function: characterization of the required calculation time.

@@

The mutation operators based on variable normal or variable uniform laws work very poorly and are difficult to configure. Many operators present more or less the same performance scores and require a variable amount of assessments. The convergence analysis \citep[see][]{Horton2012a} allows one to highlight three best operators: non-uniform mutation, chromosome of adaptive search radius, and multi-scale mutation. Thus, different optimizations were further performed using variants of these 3 operators (Figures \ref{fig:operator_mutation_score_atmlevel}-\ref{fig:operator_mutation_score_r4}).

@@

Next, a second level of analogy was proposed (Fig. \ref{fig:operator_mutation_score_r2}) based on moisture variables \citep[see][]{Bontron2004}. GAs had to optimize both levels of analogy simultaneously. Once again, despite the difficulty to do so, the results were better than the sequential calibration (red line on Fig. \ref{fig:operator_mutation_score_r2}). And finally, a preselection on air temperature was added instead of the fixed calendar window, as proposed by \cite{BenDaoud2016}. The results show generally higher scores (Fig. \ref{fig:operator_mutation_score_r4}), demonstrating the success of the optimizer to take advantage of this new degree of freedom, and its capacity to handle optimization of 3 analogy levels jointly. Again, the most relevant options are generally the same.

@@

The size of the population ($N$), i.e. the number of sets of AM parameters considered, has an effect on the performance of the optimization (Fig. \ref{fig:option_taillepop_score}). A bigger population leads to better results, but also to significantly longer optimizations. Indeed, the required number of evaluation, and thus the required time, is approximately proportional to the population size. The optimal size seems to depend on the complexity of the AM to optimize: a more complex AM (ie. with more degrees of freedom) requires a bigger population size. A rule of thumb based on a limited number of case studies (not shown here) is provided hereafter:

>>>>

In order to evaluate a combination of operators/options, 10 optimizations per configuration of GAs were processed. With all assessed combinations of GAs options, 21,630 optimizations were performed on an Intel Xeon based HPC cluster with 96 cpu, running Linux RedHat. Such assessment is not possible on the whole archive length, and had to be performed on a reduced period (5 years). The total resources needed for this comparison amounts to 229,539 h/cpu (26.2 years). One should remember that these intensive resources were needed in order to assess multiple combinations of GAs options in order to provide recommendations. Based on these, applying GAs to optimize AMs require some computational power, but not as much (the time needed is function of the archive length, the population number, the other GAs options, and the complexity of the AM to optimize).

The performances were characterized by four indicators: (i) mean performance score: average of the final scores of the 10 optimizations, (ii) convergence: the number of optimizations that converged to a supposed global optimum, (iii) number of generations: characterization of the convergence speed, and (iv) number of evaluations of the objective function: characterization of the required calculation time.

@@

The mutation operators based on variable normal or variable uniform laws work very poorly and are difficult to configure. Many operators present more or less the same performance scores and require a variable amount of assessments. The convergence analysis \citep[see][]{Horton2012a} allows one to highlight three best operators: non-uniform mutation, chromosome of adaptive search radius, and multi-scale mutation. Thus, different optimizations (an additional 4,950) were further performed using variants of these 3 operators (Figures \ref{fig:operator_mutation_score_atmlevel}-\ref{fig:operator_mutation_score_r4}).

@@

Next, a second level of analogy was proposed (Fig. \ref{fig:operator_mutation_score_r2}) based on moisture variables \citep[see][]{Bontron2004}. GAs had to optimize both levels of analogy (on the geopotential height first, and then on the moisture index made of relative humidity and precipitable water) simultaneously. Once again, despite the difficulty to do so, the results were better than the sequential calibration (bottom line on Fig. \ref{fig:operator_mutation_score_r2}). And finally, a preselection on air temperature was added instead of the fixed calendar window, as proposed by \cite{BenDaoud2016}. The results show generally higher scores (Fig. \ref{fig:operator_mutation_score_r4}), demonstrating the success of the optimizer to take advantage of this new degree of freedom, and its capacity to handle optimization of 3 analogy levels jointly. Again, the most relevant options are generally the same.

@@

The size of the population ($N$), i.e. the number of sets of AM parameters considered, has an effect on the performance of the optimization (Fig. \ref{fig:option_taillepop_score}). A bigger population leads to better results, but also to longer optimizations: the required number of evaluation, and thus the required time, is approximately proportional to the population size. Thus, if the population size is doubled, the time required for the optimization also nearly doubles (e.g., for a 5 years calibration period, approximately 4~h/cpu was needed when $N = 50$, and 14~h/cpu when $N = 200$). The optimal size seems to depend on the complexity of the AM to optimize: a more complex AM (ie. with more degrees of freedom) requires a bigger population size. A rule of thumb based on a limited number of case studies (not shown here) is provided hereafter:
